---
title: "About project"
output: html_document
---

```{r setup, include=TRUE, cache=TRUE, warning=FALSE, echo=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

Make sure the following packages are installed.

```{r installs, warning=FALSE, cache=TRUE, dependson='setup', include=TRUE}
library(tidyverse)
library(nycflights13)
library(gapminder)
library(Lahman)
```

To install a package in r, use the command `install.packages("packageName")`. Its not recommended to install packages through a Knitr document which would probably be compiled several times. 

Tidyverse package comes with many packages by default, which are frequently updated. To run the update for its packages, run `tidyverse_update()`

For getting help, you could use a package like **reprex**, which helps to generate reproducible code which can be copied for stackoverflow questions, or github snippets. To make a reproducible question, ensure the packages involved are listed at the top of the script, use `dput` function to help you generate the code to recreate the data, which can be added as a variable in the script for reproducibility, then clean up code so its neat. Read [R-bloggers](https://www.r-bloggers.com/), check out twitter tag - *[#rstats]*, follow prominent twitter R accounts, like **@hadleywickham**, **@statgarrett**, **@rstudiotips**, etc.
