---
title: "Visualisation Basics"
author: "Chidi"
date: "11/20/2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE)

library(tidyverse)
library(cowplot)
```

##### PS: Answers to exercises are what I think they are at the moment of writing. May have to come back to it when I think I know a better answer, and my answers may be wrong  

## Visualisation Basics

Data visualisation is one of the most important tools of data science. In R, the main (and most common) library for visualisation is the **ggplot2** package. In this section, I'll be going through important concepts for using this package.   

### Aestistics mapping with ggplot2  

An aestistic is a visual property of the objects in your plot. It includes things like size, colors of your points or lines or whatever, shapes, etc. The following shows how I'm mapping the `class` variable to the plot with a color aestistic on a scatterplot  

```{r eg1}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

With this idea of aestistics, it means we can add a different property to an n dimensional plot with given aestistic properties, like I can add the class above to be different shapes for each car class type, or map a condition to an aestistic property, etc.

To map something to an aestistic with ggplot, you'll have to add it in the aes() function of the mapping parameter in the geom function for which you want to map something to. In doing so, ggplot2 would automatically assign a unique level of the aestistic to each unique value of the variable, this is known as **scaling**. ggplot2 would also adds a legend to the side of the plot that explains which levels correspond to a value. All this is done automatically, but can still be controlled.  

Now I could have mapped the `class` variable to a size aestistic, like below  

```{r eg2}
ggplot(data = mpg, aes(x = displ, y = hwy, size = class)) + 
    geom_point()
```

In the plot above, each car `class` is mapped with its own point size as shown in the legend. However, I get the warning -  *Warning: Using size for a discrete variable is not advised.*, this is because I'm mapping an ordered variable (`class`) to an ordered aestistic - **size**, which is not a good idea, and therefore not recommended.

Take a look at the plots below

```{r eg3, warning=FALSE, echo=TRUE, cache.lazy=TRUE, message=FALSE}
if (!require("cowplot")) {
  install.packages("cowplot")
} else {
  theme_set(theme_bw(base_size = 8))
  plot_with_alpha_aes <- 
    ggplot(data = mpg, aes(x = displ, y = hwy, alpha = class)) + 
    geom_point()
  
  plot_with_shape_aes <- 
    ggplot(data = mpg, aes(x = displ, y = hwy, shape = class)) + 
    geom_point()
  plot_grid(plot_with_alpha_aes, plot_with_shape_aes, 
            labels = "AUTO", ncol = 1)
}
```

In the first one (alpha), the `suv` and `subcompact` categories are both plotted with the same alpha, this is because, by default only six categories can be plotted at the same time by ggplot2 for the alpha aestistic property, same with the second one (shape), in this case, `suv` isn't even mapped to a shape aestistic. Note this when setting aestistic properties with ggplot2     

Aside from setting aestistic properties to represent some variable mapping, we can set this for just the normal plotting, to do this, we have to set it outside the `aes()` parameter but inside the geom property, or generically in the ggplot constructor parameter. For example  

```{r eg4}

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), colour = "cyan", shape = 20, size = 4)
  
```

R has 25 built in shapes that are identified by numbers 1 to 25 from which we can select for our desired shapes. You can check the documentation of points `?points` to see the different shapes

### Exercises  
##### 1. What’s gone wrong with this code? Why are the points not blue  

```{r exercise1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

I think what is wrong with the above plot is that, color should receive a factor/category represented in the mpg dataset, but instead it gets the wrong type, which isn't the effect we want. If we want to make the data points blue, then we should move the color argument out of the `aes()` funtion like below  

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

##### 2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?  

```{r}
?mpg
head(mpg)
glimpse(mpg)
```

Categorical variables in the dataset includes `manufacturer`, `model`, `trans`, `drv`, `fl`, `class`. This can be seen from the above under each column name, alternatively I can use `glimpse(mpg)` to see this.   

```{r cache=TRUE}
mpg %>%
  select(which(sapply(., function(x) typeof(x) %in% "character"))) %>%
  sapply(., table)
```

The above is the categorical variables and the number of each category in the categorical variable. Continuous variables in the dataset would be `cty`, `hwy`, `displ` and `cyl` (I think there can be any number of this as well), `year` would be a discrete variable

##### 3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?  

```{r}
if (!require("cowplot")) {
  install.packages("cowplot")
} else {
  theme_set(theme_bw(base_size = 8))
  plot_with_size <- 
    ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, size = displ))
  
  plot_with_colour <- 
    ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy, colour = cty))
  
  plot_grid(plot_with_size, plot_with_colour, 
            labels = "AUTO", ncol = 1)
}
```

From the above, if it were a categorical variable, it would have shown which category was mapped to which aestistic property (colour in this example) in the legend on the side, this is different from a continuous variable in which categories aren't clear in the plot (rightly so though as I don't have a category there). This is also some useful plot though as it can be used in some kind of context (I think), like to show some progress across data points and maybe find patterns, something like that. Note that a continuous variable cannot be mapped to a shape aestistic as the following would throw an error  

```
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = displ))
  
# Error: A continuous variable can not be mapped to shape
```

##### 4. What happens if you map the same variable to multiple aesthetics?  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class, shape = class)) +
  scale_shape_manual(values = c(0:7))
```

When mapping the same variable to multiple aestistics, it applies those aestistics to each of the categories, like from the above, I mapped class to both size and colour aestistics, so it applied different sizes to each of the class category, and different colours to each of the category. Doing this doesn't really any value to the plot than if I had used one aestistic property for the variable.  

##### 5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

From the documentation, stroke aestistic is used to modify the border width of shapes that has border like shape 21. Example  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), 
             shape = 1, 
             stroke = 3, 
             colour = "red",
             fill = "white")
```

##### 6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < mean(displ)))
```

It tries to create a variable from the expression and applies that aestistic property to data points of the expressions result. In the example above, it creates a categorical variable of either true or false, and applies the aestistic property based on the true or false values. The above is similar to the following  

```{r}
mpg %>%
  select(hwy, displ) %>%
  mutate(above_mean = displ > mean(displ, na.rm = T)) %>%
  ggplot(data = .) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = above_mean))
```

### Facets  

Alongside using aestistics to add additional variables to our plots, we can use facets to add additional variables to our plotting too, this is especially useful with categorical variables. Facets are like sub plots that each display one subset (based on the faceting category) of our data. With ggplot, the function for achieving this is `facet_wrap()`. We can only facet on discrete variables.  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
  facet_wrap(~ manufacturer, nrow = 2)
```

The first argument to `facet.wrap()` is a formula (this is different from an equation as here it denotes the name of a data structure). We create a formula using `~` (a tilde) followed by the name of the variable we're faceting on. To facet our plot on a combination of two variables, we add `facet_grid()` to our plot, unlike `facet_wrap()`, the formula for `facet_grid` is, first one of the variable names, followed by a tilde `~`, followed by the other variable name. Example 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
  facet_grid(drv ~ cyl)
```

The above plots `hwy` against `displ` in facets of `drv` by `cyl`. Facets without data points are still shown. I can achieve a similar plot `facet.wrap`, but it is organised differently, like below  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
  facet_wrap(drv ~ cyl)
```

Note from the above plot that there's no facet 4-5 (`drv` ~ `cyl`), because there're no data points for it, hence the reason why its not there in `facet_wrap`. If I want to not facet in the rows and columns dimensions with `facet_grid`, I can use a `.` instead of the variable name (as shown in the example below). This shows that `facet_wrap` and `facet_grid` are very similar  

```{r}

if (!require("cowplot")) {
  install.packages("cowplot")
} else {
  theme_set(theme_bw(base_size = 8))
  with_grid <- ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy)) +
    facet_grid(. ~ cyl)
  
  with_wrap <- ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy)) +
    facet_wrap(~ cyl)
  
  plot_grid(with_grid, with_wrap, 
            labels = c("Facet_grid", "Facet_wrap"), ncol = 1)
}
```

From the above examples, I think the differences between `facet_wrap` and `facet_grid` has to do mainly with the formatting of the outputted plot. 

### Exercises  
##### 1. What happens if you facet on a continuous variable?
```{r cache=TRUE}
theme_set(theme_bw(base_size = 8))

plot_grid(
  ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
    facet_grid(. ~ cty),
   ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
    facet_wrap( ~ year), 
            labels = "AUTO", ncol = 1)
```

It does the faceting like it was doing so on a categorical variable. That means it groups the data by unique values of the continuous variable, and makes the plot for each of the group.  

##### 2. What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?  
It means that there are no data points to plot for that facet.  The following shows number of values in the grid of `drv ~ cyl`   
```{r}
table(mpg$drv, mpg$cyl)
```

In the above places with zero values would be empty in the plot, as there're 0 values in those groups.

##### 3. What plots does the following code make? What does . do?
For a `facet_grid`, it is used to create the formula that facets with one variable in a given axis (based on where the `.` is). The following example shows this
```{r cache = TRUE}
plot_grid(
  ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
    facet_grid(. ~ year),
   ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy, colour = class)) +
    facet_grid(year ~ .), 
            labels = "AUTO", ncol = 1)
```

Its useful to pay attention on the direction to use when doing to be able to see the plotting better.  

##### 4. Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

##### What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?  
One advantages of faceting against using aestistics is that there might be few data points to plot, and so the plot might not be of much information, which would not be the case if the data set was larger. 
Another advantage of using colour as facet instead of a colour aestistic is that, I can see the whole pattern clearly and better for each of the colour, which isn't clear with the aestistic because of noise from other points plotted on an a colour aestistic plot. The above can be an advantage and a disadvantage based on what we want to achieve, basically if all those. Another advantage would be that we may want to add another dimension, which would be by using a facet and an aestistic together.  

##### 5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?  
According to the documentation, "facet_wrap wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than facet_grid because most displays are roughly rectangular.". `nrow` and `ncol` is the rows and columns the plots should fit in. Other options are `shrink`, `labeller`, etc. `facet_grid()` doesn't have the `ncol` and `nrow` argument because it doesn't need it as the ncol and nrow is provided by the unique values of the variables in the faceting formula.  

##### 6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?  
```{r cache = TRUE}
plot_grid(
  ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy)) +
    facet_grid(. ~ manufacturer),
  ggplot(data = mpg) +
    geom_point(mapping = aes(x = displ, y = hwy)) +
    facet_grid(manufacturer ~ .), 
  labels = c("by column", "by row"),
  ncol = 1, label_size = 8)

```

Taking a look at both plots above, whats obvious is that when there're many unique values, its better to have them in the column, than in the rows because its easier to see the values and make comparisons.  

### Geometric Objects  
When plotting, we can make plots of the same variables but with different visual representation of the data, in ggplot2, we say with different **geoms**, which shortens for *geometric objects*. For example, see the plots below  

```{r warning=FALSE}
theme_set(theme_bw(base_size = 8))
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
plot_grid(ggplot(data = mpg) +
            geom_point(mapping = aes(x = displ, y = hwy)), 
          ggplot(data = mpg) +
            geom_smooth(mapping = aes(x = displ, y = hwy)),
          labels = "AUTO", ncol = 1)
```

They both plot the same variables, but use different geometric objects (geoms). Formally, a **geom** is a geometric object that a plot uses to represent data. So usually, plots are defined by the type of geom that was used to describe the data. 

An example of `geom_smooth` with different aestistic properties would be  

```{r warning=FALSE}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, colour = drv))
```
In the plot above, I'm drawing a smooth line for each `drv` category with a different color

We can use different geoms on a plot. In the example before the one above, plot A is a scatter plot, which uses a point geom, and plot B is a smooth geom, which adds a smooth line to the data to see trends.  
I can combine both together into a single plot in ggplot2;  
 
```{r warning=FALSE}
ggplot(mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(aes(linetype = drv, colour = drv))
```

From the above, we can see that, I've combined different geometric properties, as well as specific aestistics for these geoms in a single plot.  Note in the plot above how I put the `mapping` argument in the `ggplot` call and not in the `geom` function, the mapping above would act as the global argument for its `geom` functions, unless I want to use a different mapping for a geom function. This is how we make things DRY with ggplot2.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

In the example above, `geom_smooth` is using its own data, which would override that provided in `ggplot` function, but it'll still use the aes property in its `ggplot` call since that was overriden, then it draws the smooth line for that filter. Its kind of doing the same thing as the following (ignoring the false argument line, and I intentionally decided not to add the points)  

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_smooth(aes(colour = class == "subcompact"), se = FALSE)
```

ggplot2 has over 30 geoms, and even more in an extensions package found [here](https://www.ggplot2-exts.org). A good way to see ggplot2 geoms would be to use the cheatset provided. You can download it [here](http://rstudio.com/cheatsheets), or in Rstudio, click on *help > Cheatsheets > Data visualization with ggplot2* to download the cheatsheet. Remember that you can always check about a geom function like any other function using `?function_name` 


### Exercises 
##### 1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?  

```{r}
#> Line graph connects the points in the plot together
plot_grid(ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
            geom_line() + # draws the line chart.
            geom_point(shape = 21, fill = "white") +
            ggtitle("Line graph"), 
          ggplot(data = mpg, mapping = aes(x = class, y = displ)) +
            geom_boxplot() + # draws the boxplot
            ggtitle("Box-plot") +
            ylab("Engine displacement") +
            xlab("Type of car"), 
          ggplot(data = mpg, mapping = aes(hwy)) +
            geom_histogram(aes(y = ..density..), 
                           binwidth = .5, 
                           colour = "white",
                           fill = "black") + # geom_histogram plots a histogram
            ggtitle("Histogram"),
          ggplot(data = economics, mapping = aes(x = date, y = unemploy)) +
            geom_area() +  # makes an area plot
            ggtitle("Area chart") +
            ylab("Unemployed"),
          ggplot(data = mpg, aes(hwy)) +
            geom_area(stat = "bin") + # also makes an area plot
            ggtitle("Area chart"),
          ncol = 1)
```

##### 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.  

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

Haha, I predicted well. 

##### 3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?  

show.legend = FALSE would hide the legend that shows on the right of a plot with an aestistic property. By deafult its true, hence why we see the legend when we don't specify this argument. 

##### 4. What does the se argument to geom_smooth() do?  

`se` would remove the confidence interval drawn in around the smooth line in `geom_smooth` 

##### 5. Will these two graphs look different? Why/why not?  

```{r}
first <- ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

second <- ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

plot_grid(first, second)
```

They'll not look different because they'll both using the same data and plotting the same things, the first plot is better because it helps us avoid repeating ourselves as the second one shows.  

##### 6. Recreate the R code necessary to generate the following graphs.  

```{r}
plot_grid(ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
            geom_point() +
            geom_smooth(se = FALSE),
          ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
            geom_point() +
            geom_smooth(aes(group = drv), se = FALSE),
          ggplot(data = mpg, mapping = aes(x = displ, y = hwy, colour = drv)) +
            geom_point() +
            geom_smooth(se = FALSE),
          ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
            geom_point(aes(colour = drv)) +
            geom_smooth(se = FALSE),
          ncol = 2)
```