---
title: "Visualisation Basics II"
author: "Chidi"
date: "12/14/2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)
```

### Introduction  

**Statistical transformations** (stats) allows us plot summaries of the data, and are useful for visualisations like the histogram.  **Position adjustments** helps handle overlapping objects. **Coordinate systems** control how x and y map to position on the plot.

### Statistical Transformations  

```{r include = FALSE, warning = FALSE}
library(tidyverse)
library(nycflights13)
library(cowplot)
library(maps)
library(mapproj)
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

To make a plot, charts like histogram, bar charts, frequency polygons, etc bin your data and then plot bin counts. Smooters fit a model on the data and then plot predictions from the model. Boxplots compute a robust summary of the distribution which it displays in a specially formatted box.  For each geom we use, we can check the `stat` it uses using the `stat` argument of the geom (check the geom's doc - `?geom_bar`), and we can set for the one we want. The *Computed variables* section for these stat functions tells how it computes the stats it uses, and provides good pointers on the stats used by the geom function. We can create the geom functions of statistical transformation graphs using their direct stat functions, for example, I can recreate the bar chat above using the `stat_count`, note that the one I made above used the `count` stat, I could still use the `prop` stat, which would compute groupwise proportions  

```{r}
ggplot(diamonds) +
  stat_count(mapping = aes(x = cut))
```

This works because every geom has a default stat, and every stat has a default geom. This means we can plot graphs with statistical transformations without having to worry about the stats used, however, there're reasons why it may be important to specify the kind of stat we want;

* We might want to override the default stat or geom. for example, say I want to create a bar chart, where I have a category and the size of the category (note, not the computed frequency)  

```{r}
demo <- tribble(
  ~cut,        ~freq,
  "Fair",      1610,
  "Good",      4960,
  "Very Good", 12082,
  "Premium",   13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

The above plot shows a bar chart, however, note that the `cut` axis is the one we provide, this is the most common way bar charts are thought of, unlike the earlier example where the count was computed and we didn't have to provide `freq`  

* You might want to override the default mapping from transformed variables to aestistics. For example, I might want a bar chart to show proportions instead of count. Example;  

```{r}
ggplot(diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

* I might want to draw more attention to the statistical transformation. For example, I might want to use `stat_summary()`, which would summarise the y values for each unique x value.  

```{r}
ggplot(diamonds) +
  stat_summary(mapping = aes(x = cut, y = depth),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median, geom = "pointrange")

diamonds %>%
  group_by(cut) %>%
  summarise(minimum = min(depth, na.rm = TRUE),
            maxinum = max(depth, na.rm = TRUE),
            average = median(depth, na.rm = TRUE))
```

What the plot does is plot the summary I made below for each cut group. That is awesome. ggplot2 provides over 20 `stat` functions we can use. Use the data visualisation cheatsheet to also learn more, also check the doc for each to learn more, eg `?stat_bin`  

### Exercises  

##### 1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?  

The default geom associated with `stat_summary()` is `pointrange`. To rewrite the plot above.  

```{r}
ggplot(diamonds, mapping = aes(cut, depth)) +
  geom_pointrange(stat = "summary",
                  fun.ymin = min, 
                  fun.ymax = max,
                  fun.y = mean)
```

##### 2. What does geom_col() do? How is it different to geom_bar()?  

```{r}
diamonds %>%
  group_by(cut) %>%
  summarise(
    count = n(),
    minimum = min(depth, na.rm = TRUE),
    maxinum = max(depth, na.rm = TRUE),
    average = median(depth, na.rm = TRUE)) %>%
  print(.) %>% 
  ggplot(., aes(x = cut, y = average)) +
  geom_col()
```

The `geom_col` creates a bar chart with the x and y variables we provide. Unlike `geom_bar` whose y axis is a statistical transformation of the number of cases of x, in `geom_col` the height of the bars is the actual values in the dataset. 

##### 3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?  

See [ggplot2 documentation](http://ggplot2.tidyverse.org/reference/) 

##### 4. What variables does stat_smooth() compute? What parameters control its behaviour?

```{r}
ggplot(diamonds, mapping = aes(carat, price)) +
  geom_point() +
  stat_smooth()
```

Variables important for `stat_smooth` are the same `geom_smooth`, these includes, x, y, ymin, ymax, se. parameters that control its behavior are `method`.  

##### 5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?  

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```

I think the group helps us know how the x values are grouped relative to the y value, kind of helps us add some dimension to the plot. So by setting the group to be 1, we say we want one bar per category, by setting the group to a categorical variable, we're saying we want that number of group categories on the x axis for each of the category. The intended plot may be; 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = color), position = "dodge")
```
### Position Adjustments  

```{r}
plot_grid(ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut)),
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut)), labels = "AUTO", nrow = 2)
```

From the above, I am using the same variable (cut) for different aestestics. However, this changes when I use a different variable for different aestistics in a bar chart.  

```{r}
plot_grid(
  ggplot(data = diamonds) + 
    geom_bar(mapping = aes(x = cut, fill = clarity)),
  ggplot(data = diamonds) +
    geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge"),
  labels = "AUTO", ncol = 1)
```

Note the difference between both graphs, its the "position" argument. The first one uses a position equal to "stack", and the second one uses a "dodge" argument, others includes "identity" and "fill".    
The position parameter is for position adjustment of the geom objects. *identity* position would place each object exactly where it falls in the context of the graph. Its not very useful for bars. To see the positioning. To see "identity" positions;

```{r}
plot_grid(ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
            geom_bar(position = "identity", alpha = 1/4),
          ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) +
            geom_bar(position = "identity", fill = NA),
          nrow = 2)
```

A bar chart of identity positions is best visualised by setting an alpha for the colours, or by setting fill colour = NA (as seen from the plots above). Identity position adjustments are more useful when working with 2D geoms.

*fill* positions works like a *stack* positioning as well, but this time, the height of each bars are all equal, but each of the categories take different proportions of each bar. Example

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
            geom_bar(position = "fill")
```

Another useful position is the *jitter*, which is a useful position parameter for scatterplots. Example, note these plots   

```{r}
plot_grid(
  ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
    ggtitle("Without Jittering"),
  ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter") +
    ggtitle("With Jittering")
)
```

In the first plot suffers from **overplotting**, which is a scenario where data points are overlaid on each other so that not all points are seen, which would be the case in a dataset with large data points. Jittering position adjustment helps us fix this issue by adding some random noise to each point, this spreads the data because no two points are likely to have the same amount of random noise. ggplot's shorthand for `geom_point(position = "jitter")` is `geom_jitter()`  

```{r}
ggplot(mpg) +
  geom_jitter(mapping = aes(x = displ, y = hwy))
```

### Exercises  

##### 1. What is the problem with this plot? How could you improve it? 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

Problem with it I think is that it sufferes from overplotting, which I'll fix by adding a jitter 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point( position = "jitter")
```

##### 2. What parameters to geom_jitter() control the amount of jittering?  

The `width` and `height` can control the amount of jittering horizontally or vertically. For example 

```{r}
plot_grid(ggplot(data = mpg) +
            geom_jitter(mapping = aes(x = cty, y = hwy)),
          ggplot(data = mpg) +
            geom_jitter(mapping = aes(x = cty, y = hwy), width = 10, height = 5),
          ncol = 1)
```

##### 3. Compare and contrast geom_jitter() with geom_count().

```{r}
plot_grid(
  ggplot(data = mpg) +
            geom_point(mapping = aes(x = cty, y = hwy)),
  ggplot(data = mpg) +
            geom_jitter(mapping = aes(x = cty, y = hwy)),
  ggplot(data = mpg) +
            geom_count(mapping = aes(x = cty, y = hwy), position = "jitter"),
          ncol = 1)
```

`geom_count` is another approach for fixing overplotting. It counts the number of variables at a given overplotting point (that would have been jittered), then maps the point count to the point areas, and provides an explanation of this mapping as a legend to the plot.  Its relationship to `geom_jitter` therefore has to do with fixing overplotting, both with their unique approach to this.  

##### 4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

The default position adjustment of `geom_boxplot()` is *"dodge"*. This can be found in the documentation.  

### Coordinate systems  

ggplot2 has several coordinate systems. The most common being the cartesian coordinate system, where the x and y positions act independently to determine the location of each point.  Other coordinate systems that are occassionally useful includes;  

* `coord_flip()`: This switches the x and y axis, which can be useful especially when we have long variable names. For example;

```{r}
plot_grid(
  ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
    geom_boxplot(),
  ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
    geom_boxplot() + 
    coord_flip())
```

In the above example, note the names of the label for the x axis on the first plot, and y axis on the second plot (due to flipping the coordinates)

* `coord_quickmap()`: Sets the aspect ratio correctly for maps. Which can be very useful when plotting spatial data with ggplot2.

```{r}
nz <- map_data("nz")
plot_grid(
  ggplot(nz, aes(long, lat, group = group)) + 
    geom_polygon(fill = "white", colour = "black"),
  ggplot(nz, aes(long, lat, group = group)) + 
    geom_polygon(fill = "white", colour = "black") +
    coord_quickmap()
)
```

* `coord_polar`: This reveals an interesting connection beteen a barchart and a coxcomb chart.  

```{r}
bar_chart <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

plot_grid(bar_chart + coord_flip(), bar_chart + coord_polar())
```

### Exercises  

##### 1. Turn a stacked bar chart into a pie chart using coord_polar().

```{r}
stacked_barchart <- ggplot(diamonds, mapping = aes(x = factor(1), fill = cut)) +
  geom_bar(width = 1)
stacked_barchart
stacked_barchart + coord_polar(theta = "y")

stacked_multiple <- ggplot(diamonds, mapping = aes(x = clarity, fill = cut)) +
  geom_bar(position = "fill")
stacked_multiple
stacked_multiple + coord_polar()
stacked_multiple + coord_polar(theta = "y")
```

I can play with variations of the above to achieve different variations, in particular changing the position argument type of `geom_bar`, adding or removing `theta = "y"` argument.  

##### 2. What does labs() do? Read the documentation.

`labs` is used to modify a ggplot axis, legend, or plot labels. Like the label of the x and y axis, the legend, plot title and subtitle. 

##### 3. What’s the difference between coord_quickmap() and coord_map()?

`?coord_map`  

##### 4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
no_coord <- ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() 

no_coord
no_coord +
  coord_fixed()
```

The plot tells me there's some linear relationship between `highway` and `city` variables in the mpg dataset. `coord_fixed` sets an aspect ratio for the x and y axis of the plot. The `geom_abline` adds a slope to the plot.

