---
title: "Ggplot2 Labels"
author: "Chidi"
date: "12/17/2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)
```

### Introduction                                                                              

Labels helps improve our visualisations so they're more readable and understandable. 

```{r warning=FALSE, echo=TRUE}
library(tidyverse)
library(nycflights13)
library(patchwork)
```

With ggplot2, we can add labels using the `labs()` function.  

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point( aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size",
       subtitle = "Two seaters (sports cars) are an exception because of their light weight",
       caption = "Data from Fueleconomy.gov",
       x = "Engine displacement",
       y = "Highway fuel economy",
       colour = "Car type")
```

I can also use mathematical equations as labels, to do this, I'll just need to switch from "" to `quote()`. Check `?plotmath` for more  

```{r}
df <- tibble(
  x = runif(10),
  y = runif(10)
)

ggplot(df, aes(x, y)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + beta + frac(delta, theta))
  )
```

### Exercises  

##### 1. Create one plot on the fuel economy data with customised title, subtitle, caption, x, y, and colour labels.  

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_jitter(aes(color = fl)) +
  labs(
    title = "Fuel efficiency decreases with engine size",
    subtitle = "Grouped by fuel type",
    caption = "Fuel economy data (fueleconomy.gov)",
    y = "Highway miles per gallon",
    x = "Engine displacement",
    colour = "Fuel type"
  )
```

##### 2. The geom_smooth() is somewhat misleading because the hwy for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_jitter(aes(color = class)) +
  labs(
    title = "Fuel efficiency decreases with engine size",
    subtitle = "Grouped by fuel type",
    caption = "Fuel economy data (fueleconomy.gov)",
    y = "Highway miles per gallon",
    x = "Engine size",
    colour = "Car type") +
  geom_smooth()
```

Note that the light weight sports cars referred here are the *2seater* car types. To fix this, I'll fit a smooth geom group by group, or I'll remove the *2seater* cars from the data set and fit a model.  

```{r warning=FALSE}
ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_jitter() +
  labs(
    title = "Fuel efficiency decreases with engine size",
    y = "Highway miles per gallon",
    x = "Engine size",
    colour = "Car type") +
  geom_smooth(aes(group = class), se = FALSE, method = "lm") +
ggplot(filter(mpg,  class != "2seater"), aes(displ, hwy)) +
  geom_jitter(aes(colour = class)) +
  labs(
    title = "Fuel efficiency decreases with engine size",
    y = "Highway miles per gallon",
    x = "Engine size",
    colour = "Car type") +
  geom_smooth(se = FALSE, method = "lm")
```

##### 3. Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.
