---
title: "Other Single Table Verbs"
author: "Chidi"
date: "12/18/2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)
```

### Introduction

Aside from `mutate`, `summarise`, `filter`, `group_by` (see my *04_manipulation_basics.Rmd* file), there're other verbs that are very important to know for a start and have in one's toolbox. 

```{r warning = FALSE, echo = TRUE}
library(tidyverse)
library(nycflights13)
```

### Select  

Most timess, we work with datasets that have a large number of variables, which may or may not be necessary for our purpose. `select()`, as the name suggests, lets us select only the variables we need for our purpose. In dplyr, selects also has various helper functions that improves how efficient we can select, like `starts_with(x)`, `ends_with(x)`, `contains(x)`, `matches(x.y)`, and selections like `a:e`, which would select all variables between `a` to `e`. Example;

```{r}
flights %>%
  select(year:day, ends_with("delay"))
```

To remove variables from a selection, add `-` in front of the expression.  

```{r}
flights %>%
  select(-ends_with("delay"), -(year:day), -contains("dep"))
```

Say I want to move variables to the beginning, I'll use the `everything()` verb  

```{r}
flights %>%
  select(contains("dep"), everything())
```

### Renaming  

`rename`, as the name suggests, is used to rename variables in the format `rename(new_name = old_name)`

```{r}
flights %>%
  rename(arrival_delay = arr_delay) 
```

We can also rename with `select`, which can save us a line in case we want to select and rename too  

```{r}
flights %>%
  select(year:day, tail_num = tailnum)
```

### Transmute  

`transmute` is like `mutate`, being that it lets us create new variables from functions of the old one, the difference here being that, while `mutate` would return a dataset composed of variables of old one and the newly created ones, `transmute` would return a dataframe of only the newly created variables.  

```{r}
flights %>%
  transmute(departed_earlier = dep_delay < 0)
```

### Arrange  

`arrange` allows us change the order of rows. To put a column in descending order, use `desc()`  

```{r}
flights %>%
  arrange(desc(dep_delay))

flights %>%
  arrange(year, month, day)
```

### Distinct  

`distinct` removes duplicates from a dataset, the result is then ordered by first occurence in the dataset  

```{r}
flights %>%
  distinct(carrier, flight)
```

### Sample  

When working with large datasets, we may want to take a random sample of the data set, this is what `sample_n()` (selects the same number of observations for each group), `sample_frac()` (selects the same proportion) does. 

```{r}
flight_groups <- flights %>%
  group_by(dest) %>%
  filter(n() > 1000)

# 100 rows
flight_groups %>%
  sample_n(100) %>%
  filter(dest == "IAH") %>%
  dim

flight_groups %>%
  sample_frac(0.01) %>%
  filter(dest == "IAH") %>%
  dim
```

### Slice rows

`slice()` allows us pick observations by positions or by group. `head()` and `tail()` are slicing types for just the begining and the ending of the dataset. 

```{r}
flights %>%
  group_by(dest) %>%
  slice(1:5)

flights %>%
  head

flights %>%
  tail
```

