---
title: "Exploratory Data Analysis 1"
author: "Chidi"
date: "12/29/2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)
```

### Introduction  

EDA is a useful technique that lets us explore our data using visialisations, transformations, and modelling. EDA helps us   

* Generate questions about our data.  
* Search for answers by visualisations, transforming, and modelling our data.  
* Use what we learn to refine our questions and/or generate new questions.  

EDA is an important part of any analysis, even if we've been given questions to explore and commnicate the result, we'll still need to investigate the quality of the data. Data cleaning is an application of EDA, because we need to investigate whether the data meets expectations. Data cleaning usually involves all aspects of EDA (transformation, visualisations and modelling)  

```{r warning = FALSE, echo = TRUE, message = FALSE}
library(nycflights13)
library(tidyverse)
library(patchwork)
```

### 1D EDA - `dylyr::count`

When doing EDA, its important to keep our mind open to what the data has to tell. Before we can start combining variables together, its important to look at a variable very well first. Dpylr's `count()` can be a handy function when doing this  

```{r}
flights %>% 
  count(carrier)
```

The above is shorthand for `flights %>% group_by(carrier) %>% summarise(n = n())`. `count` has two parameters; these includes `sort` and `wt`. `wt` can be useful as it performs the sum on another variable grouped by the initial variable;

```{r}
flights %>%
  count(carrier, wt = distance)
```

What the above does is sum all distance for each carrier group. Without the `wt` argument, *n* would be the count of each carrier group.

We can also apply `count()` on an expression, for example;

```{r}
flights %>%
  count(is.na(dep_delay))

flights %>%
  count(dep_missing = is.na(dep_time),
        arr_missing = is.na(arr_time))
```

This can be a useful technique to make quick summaries on variables as seen above. Also, we can combine `count()` with `cut_` functions from `ggplot2` to compute histograms;

```{r}
# 5 bins with equal widths
flights %>%
  count(cut_interval(arr_delay, 5))

# 5 bins with approximately equal number of points
flights %>%
  count(cut_number(arr_delay, 5))

# Hourly bins 
flights %>%
  count(cut_width(arr_delay, 60, boundary = 0))
```

### Questions 

> “Far better an approximate answer to the right question, which is often vague, than an exact answer to the wrong question, which can always be made precise.” — John Tukey  

The goal when doing EDA should be to develop a better understanding of the data. To do this, the best way is to use questions as a guide for investigating the data. When we ask questions, the question focuses our attention on specific portions of the data, and from those portions of the data and the question, we can decide on the transformations, graphs or models to make.  

The key to asking quality questions is by asking lots of questions. There're no rules on which questions to ask, however, two kinds of questions that are always useful for making discoveries within our data can be generalised as;  

> 1. What type of variation occurs within my variables?  
> 2. What kind of covariation occurs within my variables?

### Variation

Variation is the likelihood of the values of a variable to change from measurement to measurement. A value is the state of a variable when it is measured. We can see variation in real life, like when we measure any continuous variable (the value), we can get different results. Even categorical variables can have variations, like if we measure the eye colours of different people. Each variable has its own pattern of variation, which can reveal something interesting. The best way to understand the pattern of variation in a dataset is to visualise the distribution of the variable's values. 

#### Visualising distributions  

How we visualise the distribution of a variable depends on whether the variable is categorical or continuous. **Categorical variables** only take a small set of values. In R, they're usually saved as factors or character vectors. We usually use a barchat to visualise the distribution of a categorical variable.  

```{r}
ggplot(diamonds) +
  geom_bar(mapping = aes(x = cut))
```

A variable is **continuous** if it can take an infinite set of ordered values. Examples are numbers and datetime. We'll usually use a histogram to visualise continuous variables;

```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.2)
```

Note that we can use our `count` variables to compute things that are plotted here for the categorical variable and continuous variables;

```{r}
diamonds %>%
  count(cut)

diamonds %>%
  count(cut_width(carat, 0.2))
```

A histogram divides the x axis into equally spaced bins and uses the height of each bar to show the number of observations in that bin. Its a good practice to always practice with different `binwidths` when working with histograms as they can reveal different distribution patterns. 

If we wish to overlay multiple histograms on the same plot, its recommended to use `geom_freqpoly()` instead of `geom_histogram()`. `geom_freqpoly` does the same calculation as `geom_histogram`, but draws its own with lines instead of bars.  

```{r}
# local function just to allow me compose plots together with the same dataframe, since I'm piping
drawplots <- function(df) {
  ggplot(df) +
    geom_histogram(mapping = aes(x = carat, fill = cut), position = "dodge", binwidth = 0.1) +
  ggplot(df) +
    geom_freqpoly(mapping = aes(x = carat, colour = cut), binwidth = 0.1) +
  plot_layout(ncol = 1)
}

diamonds %>%
  filter(carat < 3) %>%
  drawplots(.)
```

The key to asking good follow up questions is to use our curiousity, and our skepticism (like how could what I'm seeing be misleading?). 
When we look at histograms and barchats, we can start by asking ourselves, what values are most common and why? Which ones are rare and why? Is this observation expected? Are there unusual patterns? And what could be the reason for that? Clusters usually suggests subgroups exists in our data, and to understand subgroups, we ask ourselves questions to know how the observations within each cluster relates to each other, how the observations between separate clusters differs from each other? Why? What variables can help explain/describe these clusters and the observed differences? Why might the appearance of clusters be misleading?

### Unusual Observations  

Outliers are data points that are unusual, points that doesn't fit the pattern. Several things can cause outliers, like entry errors, or an important suggestion from the normal pattern which may be useful. In general, its good to look out for outliers, and discern what could be the cause of the outlier to know how best to handle them. Its good practice to repeat one's analysis with and without outliers. If they have minimal effect and we can't figure out why they're there, its reasonable to replace them with missing variables, else, we have to keep them and figure out why they're there. Sometimes, especially when we have lots of data points to plot for a variable, like in a histogram, outliers may not be seen easily, so we have to zoom in on it. We can do this using coord_cartesian(). Example;

```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  ggtitle("Main plot") +
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = 0:50) +
  ggtitle("Zoomed in on the Y-axis") +
plot_layout(ncol = 2)
```

Notice the heights on both plots above, the one on the left is 0 to 120k, the one on the right is 0 to 60. Now I can see the values of y with smaller values lies around 30's and 50's. I can do the same for the y axis.

# Exercises  

##### 1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
# ======= What is the distribution of the x variable =========

ggplot(data = diamonds, mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.2) +
  ggtitle(label = "A") + 
ggplot(data = diamonds, mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.2) +
  coord_cartesian(ylim = c(0,10), xlim = c(-0.75, 1)) + 
  ggtitle(label = "B") + 
diamonds %>%
  select(x, y, z) %>%
  filter(x > 0) %>%
  ggplot(mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.2) +
  coord_cartesian(xlim = c(8.9, 11), ylim = c(0:20)) + 
  ggtitle(label = "D") + 
diamonds %>%
  select(x, y, z) %>%
  filter(x > 0) %>%
  ggplot(mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.2) +
  ggtitle(label = "C") + 
plot_layout(nrow = 2)
```

Looking at the distribution of the `x` variable, from the plot on the left, I see that some points are outliers, like there seem to be few points around 0, on the right, I zoomed into this area and saw that about 8 data points have x of zero, so what could be their possible `y` and `z` values  

```{r}
diamonds %>%
  select(x, y, z) %>%
  filter(x <= 0) 
```

Looking at the section of the plot above, I see that those variables are usually 0, which means they were non existent, so those data points are probably best removed, one point, has a y value of 6.62, but I think this can also be removed because its x and y are 0. So this means, I can remove those points from the plot, as seen in plot C above. Same thing can be said for the bins on the far right (between 9 and 11), which I zoomed into in plot D above, values from 9.5 to ~11 seems like an error as they're too far off,

```{r}
diamonds %>%
  select(x, y, z) %>%
  filter(x >= 9.5) 
```

But looking at the table above, I think they're valid points and very much part of the data. However, I'm curious as to why there're so few of them, I think its because they're bigger, and the bigger, the rarer, which is kind of what the x distribution shows, that as diamonds with longer widths are rare.  

To look at all variables x , y and z distributions, I'll exempt the ones I know shouldn't be there  

```{r}
# I could optimise how I use this with patchwork package, instead of repeating this, could be good for a # contribution 
plt <- function(df) {
  df %>%
    gather(dimension, dimension_value, x:z) %>%
    ggplot(mapping = aes(x = dimension_value, colour = dimension)) +
    geom_freqpoly(binwidth = 0.2)
}
diamonds

(diamonds %>% 
  gather(dimension, dimension_value, x:z))

diamonds %>%
  select(x, y, z) %>%
  filter(x > 0 | y > 0) %>%
  plt() 
```

There seems to be outliers here, first, I notice my initial observation, that the bigger the dimension, the rarer, that means those ones that are big here, indicated by the flat line in the plot needs further checking. Another thing I notice is that I can't seem to find the line for x, the red line, it's completely masked by the y variable, suggestion that the x and y are usually almost the same, this seems to suggest to me that the points are usually the same for the x and y variables, except on few occassions. If that is the case, why are there data points of y greater than x, like I see that the maximum x variable is 10.74 and y is 58.9, if majority of the data sets seems to suggest that x is always the same of close to y, then y the difference? Are those ones better diamond? More expensive, or what? Also, does dimension determine price?

```{r cache = TRUE}
## Does dimension determine price
diamonds_flattened <- diamonds %>%
  filter(x > 0) %>% # I've already established that x < 0 is wrong
  gather(dimen_categories, dimension_values, x:z)

diamonds_no_outliers <- diamonds %>%
  filter(y < 30, z < 30) %>%
  gather(dimen_categories, dimension_values, x:z)

diamonds_flattened %>%
  select(price, starts_with("dimen")) %>%
  ggplot(mapping = aes(x = dimension_values, y = price, colour = dimen_categories)) +
  geom_point() + 
  ggtitle("With outliers") +
diamonds_no_outliers %>%
  select(price, starts_with("dimen")) %>%
  ggplot(mapping = aes(x = dimension_values, y = price, colour = dimen_categories)) +
  geom_jitter() +
  ggtitle("Without outliers") +
  scale_x_discrete(limits = seq(0, 12, 2.5)) +
plot_layout(nrow = 2)
```

From the above plot, I see there're values with bigger dimensions but still small price, and one low, and its not enough to suggest that those values are far off because they're more expensive, suggesting they might have to be removed as errors, which results in the plot without outliers above. I see from the plots generally that the bigger the width or height or depth, the bigger the price. which is understandable, there are some variables of z which are 0, I think these should also be removed as well as a diamond must have depth.  

##### 2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)  

```{r cache = TRUE}
diamonds %>%
  ggplot(mapping = aes(x = price)) +
  geom_histogram() +
  ggtitle("With no binwidth set") +
diamonds %>%
  ggplot(mapping = aes(x = price)) +
  geom_histogram(binwidth = 70) +
  ggtitle("With binwidth = 70") +
diamonds %>%
  ggplot(mapping = aes(x = price)) +
  geom_histogram(binwidth = 50) +
  ggtitle("With binwidth = 50") +
diamonds %>%
  ggplot(mapping = aes(x = price)) +
  geom_histogram(binwidth = 30) +
  ggtitle("Zoomed into vacant x bins") + 
  coord_cartesian(ylim = c(0,500), xlim = c(1000, 2000)) +
plot_layout(nrow = 2)
```

I see that there're no diamonds that is 1500 dollars, which I didn't see with some binwidths. 

##### 3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r cache = TRUE}
## I Don't really have an answer as to the cause of the difference in the question
diamonds %>%
  summarise(carat_point99 = sum(carat == 0.99),
            carat_1 = sum(carat == 1))
diamonds %>%
  mutate(is_1_carat = (carat == 1 | carat == 0.99)) %>%
  select(is_1_carat, carat) %>%
  ggplot(mapping = aes(x = carat, fill = is_1_carat)) +
  geom_histogram(binwidth = 0.05) +
  scale_color_viridis_d(0.5)
```

##### 4 Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

```{r}
ggplot(data = diamonds, mapping = aes(x)) +
  geom_histogram() +
  xlim(c(9, 11)) +
ggplot(data = diamonds, mapping = aes(x)) +
  geom_histogram() +
  coord_cartesian(xlim = c(9, 11)) +
plot_layout(nrow = 2)
```

Looking the the two functions above, I notice that, with `coord_cartesian`, setting only `xlim` or `ylim` doesn't zoom appropriately with the corresponding x or y axis, unlike using either `xlim()` or `ylim()`. So `coord_cartesian` should be while providing both the `xlim` and `ylim` parameters accordingly.  

