---
title: "Data structure basics"
author: "Chidi"
date: "1/4/2018"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)
```

### Introduction  

Vectors are related data organised in data frames and tibbles. They can be *atomic* or *augumented*. 

```{r warning = FALSE, echo = TRUE, message = FALSE}
library(tidyverse)
```

#### Atomic Vectors  

Atomic vectors are the simplest data structures upon which other structures are built in R. There are 4 atomic vectors important to data analysis in R  

* logical  - `TRUE` or `FALSE`, `<lgl>` in tibble 
* integer - integers, `<int>` in tibble  
* double - real numbers, `<dbl>` in tibble  
* character - strings, `<chr>` in tibble   

Vectors can contain *NA* values, which are known as missing values. *integer* and *double* vectors are collectively known as *numeric vectors*. We can create atomic vectors using `c()` function. 

```{r}
logical_vector <- c(TRUE, FALSE, FALSE)
integer_vector <- c(1L, 2L, 3L, NA)
double_vector <- c(1, 2, 3)
double_vector2 <- c(1.5, 0.2, 3, NA)
character_vector <- c("chi", "di", "eb", "re")
typeof(logical_vector)
typeof(integer_vector)
typeof(double_vector)
typeof(double_vector2)
typeof(character_vector)
```

#### Augumented Vectors  

Augumented vectors are atomic vectors with additional metadata. There're 4 important `augumented` vectors;  

* factors `<fct>` - These are represented by categorical characters, are there're a fixed number of these characters, known together as the level of the factor.  
* ordered factors `<ord>` - These are like factors above, but the levels have an intrinsic ordering, like we can say this one is greater or less than the other, etc.    
* dates `<dt>` - A date record  
* date-time `<dttm>` - Also known as **POSIXct**, represents a date and time vector  

Related vectors (augumented or atomic) are can be grouped together to form a data frame or tibble.  

```{r}
(my_tibble <- tibble(x = c(1,9,5),
                    y = c(TRUE, TRUE, FALSE),
                    z = c("apple", "pear", "banana")))

(my_tibble2 <- tribble( # notice the name "tribble" not "tibble", but it's used to create a "tibble"
  ~x, ~y, ~z,
  1, TRUE, "apple",
  9, TRUE, "pear",
  5, FALSE, "banana"
))
```

To get the variable types of a data set, we can use `purrr::map_chr()` to apply a type sum (type summary) to each variable.  

```{r}
type_sum(diamonds)
map_chr(diamonds, type_sum)
```


