---
title: "Missing Values"
author: "Chidi"
date: "1/15/2018"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)
```


### Introduction  

```{r warning = FALSE, echo = TRUE, message = FALSE}
library(tidyverse)
library(nycflights13)
```

R uses **NA**(Not Available) to represent unknown but potentially measurable values. Almost any operation involving NA's would return NA. 

```{r}
NA > 5
10 == NA
NA + 10
NA / 2
NA == NA
```

To understand why this is so;  

```{r}
marys_age <- NA # basically saying we don't know
janes_age <- NA # same as above
marys_age == janes_age # obviously we don't know
marys_age == marys_age # same thing, we don't don't know (R doesn't use reference based assignment :) )
c = 4
d = c
c = 7
d
c
```

To check for NA values, 

```{r}
is.na(marys_age)

is.na(c(32, NA)) ## checks for each item in the vector if it is NA

```

Filtering operation from *dplyr* package returns only items in what is being filtered that evaluates to true and excludes values that are false or NA  

```{r}
mylist = tibble(x = c(2,56,1,5,9,0))
filter(mylist, x %% 2 == 0)
```

### 5.2.4 Exercises 

```{r}
path = paste0(getwd(), "/04_manipualtion_basics.Rmd" )
```

I solved the the exercises for this section in  [04_manipulation_basics.Rmd](04_manipualtion_basics.Rmd).  



