---
title: "About project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Resurrect R

This project is about me bring back my data science with R skill. I am passionate about data science, and applying the knowledge in both research and production. So my aim would be data science with R and Kotlin. Python is also there as well, but my main focus would be R and Kotlin. So I'll have similar projects for both Kotlin and Python coming.

I'm using this [site](https://dcl-2017-04.github.io/curriculum/upcoming.html) as the curriculum for this.

## Set up

Make sure the following packages are installed

```{r include=TRUE}
install.packages(c("nycflights13", "gapminder", "Lahman", "tidyverse"))
```

Tidyverse package comes with many packages by default, which are frequently updated. To run the update for its packages, run 

```{r}
library(tidyverse) # First make sure the package is loaded
tidyverse_update() # call its update function
```

For getting help, you could use a package like **reprex**, which helps to generate reproducible code which can be copied for stackoverflow questions, or github snippets. To make a reproducible question, ensure the packages involved are listed at the top of the script, use `dput` function to help you generate the code to recreate the data, which can be added as a variable in the script for reproducibility, then clean up code so its neat. Read [R-bloggers](https://www.r-bloggers.com/), check out twitter tag - *[#rstats]*, follow prominent twitter R accounts, like **@hadleywickham**, **@statgarrett**, **@rstudiotips**, etc.
